
## Measures

### Socioeconomic Status (SES)

Measures of parent occupation were obtained for both mother and partner during child gestation. Parental occupations were then classified as unskilled(0), partly skilled (1), skilled manual (2), skilled non-manual (3), manegerial and technical (4), or professional (5). Of reported parent occupations, the highest reported occupation was used as a proxy for parent SES, to be treated as a continuous (0-5) covariate in analyses.


### Body Mass Index Z-score (BMI-Z) at Age 10

```{r bmi, include = FALSE}
# clinic <- sum(!is.na(demo_ed.f$bmi_clinic.13))
# no_clinic <- sum(is.na(demo_ed.f$bmi_clinic.13))
# best_avail <- sum(!is.na(demo_ed.f$bmi_bestavail.13))
# 
# missing_n <- 7688- best_avail
# pr_n <- best_avail-clinic
# missing_percent <- round ((7688 - best_avail)/7688*100 , 2)
# clinic_percent <-  round ((clinic)/7688*100, 2)
# pr_percent <- round((best_avail-clinic)/7688*100, 2)

```

Child BMI-Z was assessed via two indices at age 13 years. First, objective height and weight were obtained for those who completed a clinic visit at age 13 (median age at clinic visit = 13.8 years). Second, parents reported child height and weight via questionnaire (median age = 13.1). An age and sex adjusted BMI-Z score was obtained from these self-report and clinic-recorded measurments. In order to maximize accuracy, when clinic-obtained measurements were obtained, this BMI-Z score was used (n = r clinic;  clinic_percent %. BMI-Z from parent report was used when clinic measurement was not available (n = pr_n;  pr_percent%). A portion of the sample had neither measurement available (n = missing_n`; missing_percent%)

### Exercise Measures

Questions related to maladaptive exercise were self-reported at ages 14, 16, 18, and 24 years. Questions were adapted from the Youth Risk Behavior Survelliance System [@kannYouthRiskBehavior1995].

#### Exercise for Weight Loss

At ages 14, 16, and 18 years, participants were asked if they exercised to lose weight or avoid gaining weight, with response options of: No(0), Yes-Sometimes(1), and Yes-Frequently(2).

At age 24, participants were asked the frequency with which they exercised to lose weight or avoid gaining weight, with response options of: Never, \<1x/mo, 1-3x/mo, 1-4x/wk, and 5 or more times per week. For the purposes of analysis, responses at age 24 were harmonized with age 14, 16, and 18. Responses of Never were binned as "No" (0); 1x/mo - <1x/week were binned as 'Yes - Sometimes'(1), and 1x/wk or more as 'Yes - Frequently'(2). Once a week or more was chosen as the threshold for 'frequent' exercise for weight loss at age 24 to align with DSM-V diagnostic frequency thresholds for other eating disorder behaviors, such as binge eating and purging [@americanpsychiatricassociationDiagnosticStatisticalManual2013]. The primary derived outcome variable of analysis for exercise for weight loss is an ordinal variable, defined as 0 = No, 1 = Sometimes, and 2 = Frequently. 

#### Maladaptive Exercise

Issues that are associated with exercise were also reported at ages 14,16, 18, and 24 years. At all assessment points, participants reported whether exercise interferes with work/school (14,16,18), or their daily routine (24). At ages 14 and 24, participants also reported whether they exercised to lose weight even when sick or injured. At ages 16 and 18 years, participants reported whether they felt guilty about missing an exercise session -- see Table \@ref(tab:tab2) .  Responses options for exercise issues were 'No', 'yes - Sometimes', or 'Yes - Frequently'. 

```{r tab2}
# x <- tibble(age = c(14,16,18,24), Interferes = c('x','x','x','x'), Sick_Injured = c('x',' ',' ', 'x'), Guilt = c('' ,'x','x', '')) 
# 
# knitr::kable(x, caption = 'Exercise Issues Assessed at Each Age')
```

Exercise issues are deemmed to be 'present' at each time point if an individual endorses: (1) exercise interfering with work/school sometimes or more OR frequent guilt when missing an exercise session at ages 16 and 18 years and (2) exercise interfering with work/school/daily routine sometimes or more OR exercising even when sick/injured sometimes or more at ages 14 and 24 years. 

The presence of maladaptive exercise in this study was defined as exercising for weight loss **sometimes or frequently** and at least one **exercise issue** defined as  present at each time point. Frequency of exercise for weight loss as sometimes or more was chosen for this definition as (1) existing literature suggests that frequency of exercise may be less important than cognitive features in defining exercise risk in the context of eating disorders [@adkinsDoesExcessiveCompulsive2005; @mondUpdateDefinitionExcessive2006] and (2) the nature of the epidemiological sample and the concomitant goals of this paper are such that we are interested in identifying early risk indicators, which may be best captured via a broad (vs. narrow) definition of maladaptive exercise. The primary variable to define maladaptive exercise was derived as described, resulting in a dichotomous outcome (0 = Absent, 1 = Present). 


#### Exercise Gropus

Exericse groups over time are defined for descriptive analyses and transition states, replicating groups defined in @schaumbergProspectiveAssociationsDriven2022. Those who report no exericse for weight loss at each time point are placed in a 'No Exercise for Weight Loss' group, those who report exercise for weight loss at least sometimes, but do not meet criteria for maladaptive exercise are placed in an 'Exercise for Weight Loss' group, and those meeting maladaptive exercise criteria are placed in the 'Maladaptive Exercise' group.

<<<<<<< HEAD
#### Eating Disorder Behaviors
=======

### Physical Activity 

Moderate-to-vigorous physical activity (MVPA) was assessed via Actigraph^tm^ accelerometers at 4 timepoints; age 11.5, 13.5, 15.5, and 24 years old. Participants were asked to wear the accelerometers for 7 days with the expception of age 24 where participants were asked to wear the accelerometer for 4 consecutive days.  Accelerometer data was used to extract total physical activity which was measured as the average counts per minute (CPM) over the period of valid recording, 480 to 600 minutes or recording in a day.[@mattocksUseAccelerometersLarge2008] Group based trajectory modeling (GBTM)[@naginGroupbasedTrajectoryModeling2010] was used to define trajectories of MVPA from middle childhood through emerging adulthood for all participants with at least one PA assessment, using Maximum likelihood (ML) estimation with estimation maximization (EM). Following estimation, posterior class probabilities were derived and used to assign individuals to classes.

### Eating Disorder Diagnoses

Eating disorder diagnoses (AN, BN,BED, purging disorder [PD], and Other Specified Feeding and Eating Disorder (OSFED)) for adolescents were derived using questionnaire data from the Youth Risk Behavior Surveillance System questionnaire (YRBSS) using DSM-5 diagnostic criteria (excluding OSFED) [@micaliAdolescentEatingDisorders2015][@americanpsychiatricassociationDiagnosticStatisticalManual2013]. AN was assessed using both the YRBSS self-report and parental report since adolescents tend to under-report AN symptoms [@houseDiagnosingEatingDisorders2008]. Table X shows the criteria used for each diagnosis [@micaliAdolescentEatingDisorders2015]. 

Diagnoses were assessed at Waves14+, 16+, 18+, and 24+. Body mass index (BMI), which is an objective measure, was also collected at face to face assessment (median ages: 13.8, 15.5, 17.8, and 24.0) and used as a diagnostic criteria for AN. Furthermore, an underweight assignment was determined using age, gender, and BMI-specific cutoffs (based on UK reference data); corresponding to World Health Organization (WHO) grade 1 thinness [@coleBodyMassIndex2007]. Excessive exercise, which is used as criteria for an AN and OSFED diagnoses, was assigned to the adolescent if they found themselves lacking time to do schoolwork due to engaging in exercise, exercising when sick or injured, and feeling guilt when unable to exercise. 

>>>>>>> a67990961bf7e7764679d41f027318c5ed2fdd83

