

## Transitions in Exercise Groups Over Time
```{r}
load('models/transition_models')
load('data/transition_data.RData')
```

A subset of girls (N = 1169) and boys (N = 568) who completed all assessments, ages 14-24. To aid with visualizing transition intensities across exercise groups and age, Figure \@ref(fig:transition-girlplot) and Figure \@ref(fig:transition-boyplot) show the frequency of individuals amongst those those completing all assessments who transitioned from each of the three exercise groups to the other groups at each age. 

```{r transition-girlplot, fig.cap = 'Transitions Across Exercise Groups in Girls'}

demo_ed_girls$exercise_group.14 <- zap_labels(demo_ed_girls$exercise_group.14)
demo_ed_girls$exercise_group.14 <- labelled(demo_ed_girls$exercise_group.14,  c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

demo_ed_girls$exercise_group.16 <- zap_labels(demo_ed_girls$exercise_group.16)
demo_ed_girls$exercise_group.16 <- labelled(demo_ed_girls$exercise_group.16,  c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

demo_ed_girls$exercise_group.18 <- zap_labels(demo_ed_girls$exercise_group.18)
demo_ed_girls$exercise_group.18 <- labelled(demo_ed_girls$exercise_group.18,   c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

demo_ed_girls$exercise_group.24 <- zap_labels(demo_ed_girls$exercise_group.24)
demo_ed_girls$exercise_group.24 <- labelled(demo_ed_girls$exercise_group.24,  c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

ex_status_girl_plot <- transition_plot(demo_ed_girls, 'exercise_group.', 'Exercise Status Across Age in Girls', na.rm = TRUE)
ex_status_girl_plot  + 
    theme(legend.text = element_text(size = 12),
          legend.title = element_blank())
```
Amongst girls, transitions are common across all groups (no Exercise for Weight Loss, Exercise for Weight Loss, Maladaptive Exercise). Among those in the Maladaptive Exercise group, at each age, transition to Exercise for Weight Loss appears common, as does remaining in the Maladaptive Exercise group, with transition to the no-Exercise for Weight Loss group being a more rare transition. Overall, stability in a group remains common, as does transition to-and-from groups that are a single step (e.g. transition from no-Exercise for Weight Loss to Exercise for Weight Loss). 


```{r transition-boyplot, fig.cap= 'Transitions Across Exercise Groups in Boys'}

demo_ed_boys$exercise_group.14 <- zap_labels(demo_ed_boys$exercise_group.14)
demo_ed_boys$exercise_group.14 <- labelled(demo_ed_boys$exercise_group.14,  c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

demo_ed_boys$exercise_group.16 <- zap_labels(demo_ed_boys$exercise_group.16)
demo_ed_boys$exercise_group.16 <- labelled(demo_ed_boys$exercise_group.16,  c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

demo_ed_boys$exercise_group.18 <- zap_labels(demo_ed_boys$exercise_group.18)
demo_ed_boys$exercise_group.18 <- labelled(demo_ed_boys$exercise_group.18,   c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

demo_ed_boys$exercise_group.24 <- zap_labels(demo_ed_boys$exercise_group.24)
demo_ed_boys$exercise_group.24 <- labelled(demo_ed_boys$exercise_group.24,  c('No Exercise for Weight Loss' = 0, 'Exercise for Weight Loss' = 1, ' Maladaptive Exercise' = 2))

ex_status_boy_plot <- transition_plot(demo_ed_boys, 'exercise_group.', 'Exercise Status Across Age in Boys', na.rm = TRUE)
ex_status_boy_plot  + 
    theme(legend.text = element_text(size = 12),
          legend.title = element_blank())
```

Amongst boys, transitions across groups also appears common, with somewhat less stability in the Maladaptive Exercise group, as compared to girls. While there is an increase in Maladaptive Exercise at age 24, as discussed in descriptive analyses, relatively few of these men also reported Maladaptive Exercise at age 18, and a moderate portion of these men reported no-Exercise for Weight Loss at age 18.

### Girls

When examining overall transitions amongst girls, visual impressions from the transition plot is confirmed, with the transition from Maladaptive Exercise to the no-Exercise for Weight Loss group being relatively rare (n = 182). Stability in the no-Exercise for Weight Loss and Exercise for Weight Loss groups was common based on raw numbers, as was transition from the no-Exercise for Weight Loss to the Exercise for Weight Loss group (raw numbers presented in Supplemental Table \@ref(tab:state-table-girls). The transition probabilities across 1-year based on the Q-matrix allowing direct transitions across all groups with the exception of instantaneous transition from 'No Exercise for Weight Loss' to 'Maladaptive Exercise' are presented in Table \@ref(tab:pmat-girls2), with a figure depicting the observed vs. expected prevalence in each group based on this model presented in Supplemental Figure \@ref(fig:psm-plot-girls). As expected, the highest probability outcome for each group over the course of a 1-year period was staying in that group, followed by transition to a neighboring group, with transitions between 'No Exercise for Weight Loss' and 'Maladaptive Exercise' groups being the least likely transitions. The model with the additional constraint on instantaeneous transitions from 'Maladaptive Exercise' to 'No Exercise for Weight Loss' (used for covariate analyses due to convergence issues), along with covariates included in the model, is presented in Supplementary Table (\@ref(tab:pmat-girls1-covs)). 

```{r pmat-girls2}
#Markov model with covariates

ex_group_girls_2_pmat <- pmatrix.msm(ex_group_girls_2, t = 1, ci = 'normal')

estimate <- as.data.frame(ex_group_girls_2_pmat[['estimates']]) %>% 
  rownames_to_column('From') %>% 
  pivot_longer(c('No EWL', 'EWL', 'Maladaptive Ex'), names_to = 'To', values_to = 'P' ) %>% 
  mutate(P = sprintf("%.3f", as.numeric(P))) 


l <- as.data.frame(ex_group_girls_2_pmat[['L']]) %>%   
  mutate('From' = c('No EWL', 'EWL', 'Maladaptive Ex')) %>% 
  rename('No EWL' = V1) %>% 
  rename('EWL' = V2) %>% 
  rename('Maladaptive Ex' = V3) %>% 
  pivot_longer(c('No EWL', 'EWL', 'Maladaptive Ex'), names_to = 'To', values_to = 'L' ) %>% 
  mutate(L = sprintf("%.3f", as.numeric(L))) 


u <- as.data.frame(ex_group_girls_2_pmat[['U']]) %>% round(3)  %>% 
  mutate('From' = c('No EWL', 'EWL', 'Maladaptive Ex')) %>% 
  rename('No EWL' = V1) %>% 
  rename('EWL' = V2) %>% 
  rename('Maladaptive Ex' = V3) %>% 
  pivot_longer(c('No EWL', 'EWL', 'Maladaptive Ex'), names_to = 'To', values_to = 'U' ) %>% 
  mutate(U = sprintf("%.3f", as.numeric(U))) 

estimate <- cbind(estimate, l$L) 
estimate <- cbind(estimate, u$U)

estimate <- estimate %>% 
  rename('l' = `l$L`) %>% 
  rename('u' = `u$U`)

 
estimate$CI <- paste(estimate$l, estimate$u, sep = ', ')
addparens <- function(x){paste('(',x,')')}
estimate$CI <- addparens(estimate$CI)
estimate$P <- paste(estimate$P, estimate$CI, sep = ' ')

estimate <- estimate %>% 
  select (c('To', 'From', 'P')) %>% 
  pivot_wider(names_from = 'To', values_from = 'P') %>% 
  rename(' ' = From)

knitr::kable(estimate, 
             caption  = 'Estimates of Transition Probabilities Across 1-year amongst Girls: From (rows) - To (cols)')

# Note - From in Rows to in Columns
```
Note. EWL = 'Exercise for Weight Loss'. Maladaptive Ex = 'Maladaptive Exercise'

Covariates were next introduced into the model, to determine whether eating disorder cognitions, BMI, and parent SES predicted transitions between exercise groups during late adolescence. To account for multiple comparisons, a 99.5% confidence interval is presented. Overall, higher BMI Z-score at age 13 appears to relate to increased likelihood of transition from 'No Exercise for Weight Loss' to 'Exercise for Weight Loss' over the course of a year, with a one standard deviation increase in BMI being associated with a 43% increase in likelihood of transition in a given year. Fear of weight gain, though not meeting the corrected significance threshold, also indicated potential association with increased likelhood of transition from 'No Exercise for Weight Loss' to 'Exercise for Weight Loss' over the course of a year. 

```{r, label = 'GirlHRs'}
hr_1_girls <- hazard.msm(ex_group_girls_covs_1)
ti_girls_hazard <- hr_1_girls[[1]]  
bd_girls_hazard <- hr_1_girls[[2]] 
fear_wt_girls_hazard <- hr_1_girls[[3]] 
bmi_z_girls_hazard <- hr_1_girls[[4]]
ses_girls_hazard <- hr_1_girls[[5]]

dfnames <- c('ti_girls_hazard', 'bd_girls_hazard', 'fear_wt_girls_hazard', 'bmi_z_girls_hazard', 'ses_girls_hazard')
Girl_HRs <- do.call(rbind, lapply(dfnames, function(x) cbind(get(x), Covariate=x))) |> 
  as.data.frame() |> 
  rownames_to_column(var = 'transition')

Girl_HRs$L <- (as.numeric(Girl_HRs$HR)-((2.807)*((as.numeric(Girl_HRs$U)-as.numeric(Girl_HRs$HR))/1.96)))
Girl_HRs$U <- (as.numeric(Girl_HRs$HR)+((2.807)*((as.numeric(Girl_HRs$U)-as.numeric(Girl_HRs$HR))/1.96)))

Girl_HRs <- Girl_HRs %>% 
 mutate(HR = sprintf("%.3f", as.numeric(HR))) %>% 
 mutate(L = sprintf("%.3f", as.numeric(L))) %>% 
 mutate(U = sprintf("%.3f", as.numeric(U))) 


Girl_HRs$transition <- gsub('\\No.', 'No ', Girl_HRs$transition)
Girl_HRs$transition <- gsub('\\.Ex', ' Ex ', Girl_HRs$transition)
Girl_HRs$transition <- gsub('\\...', ' to ', Girl_HRs$transition)
Girl_HRs$transition <- gsub('\\..*', '', Girl_HRs$transition)
Girl_HRs$CI <- paste(Girl_HRs$L, Girl_HRs$U, sep = ', ')
addparens <- function(x){paste('(',x,')')}
Girl_HRs$CI <- addparens(Girl_HRs$CI)
Girl_HRs$HR <- paste(Girl_HRs$HR, Girl_HRs$CI, sep = ' ')


Girl_HRs <- Girl_HRs |> 
  select( - c( L, U, CI)) %>% 
  pivot_wider(names_from = c(transition), values_from = HR)
Girl_HRs['Covariate'][Girl_HRs['Covariate'] == 'ti_girls_hazard'] <- 'Thin-Ideal Internalization [14]'
Girl_HRs['Covariate'][Girl_HRs['Covariate'] == 'bd_girls_hazard'] <- 'Body Satisfaction [14]'
Girl_HRs['Covariate'][Girl_HRs['Covariate'] == 'fear_wt_girls_hazard'] <- 'Fear of Weight Gain [14]'
Girl_HRs['Covariate'][Girl_HRs['Covariate'] == 'bmi_z_girls_hazard'] <- 'BMI Z-score [13]'
Girl_HRs['Covariate'][Girl_HRs['Covariate'] == 'ses_girls_hazard'] <- 'Parent Occupation'

Girl_HRs %>% 
  kbl(caption = 'Hazard Ratios for Transitions with Baseline Covariates with Transitions between Maladaptive Ex and No EWL Constrained - Girls', format = 'html')  %>% 
  kable_paper(full_width = T)   %>% 
  column_spec(2, color = case_when(Girl_HRs$Covariate %in% c('BMI Z-score [13]') ~ 'blue', 
                                    TRUE ~'black')) 
  

  


```
Note. EWL = 'Exercise for Weight Loss'. Maladaptive Ex = 'Maladaptive Exercise'

### Boys

Raw transitions amongst boys are presented in Supplemental Table \@ref(tab:state-table-boys). Overall, there appeared to be less stability in the Maladaptive Exercise group, as compared to girls, by examination of raw numbers.The transition probabilities across 1-year based on the Q-matrix allowing direct transitions across all groups with the exception of instantaneous transition from 'No Exercise for Weight Loss' to 'Maladaptive Exercise' are presented in Table \@ref(tab:pmat-boys2), with a figure depicting the observed vs. expected prevalence in each group based on this model presented in Supplemental Figure \@ref(fig:psm-plot-boys). Stability in the 'No Exercise for Weight Loss' group was common, and stability in all groups was the most common 1-year outcome. There was a moderate probability (~16-20%) of boys in the 'Exercise for Weight Loss' group transitioning to either the 'No Exercise for Weight Loss' or 'Maladaptive Exercise' groups over a 1-year period. Similarly, boys in the 'Maladaptive Exercise' group showed a moderate probability of transitioning to either the Exercise for Weight Loss (18%) or 'No Exercise for Weight Loss' (19%) group over a 1-year period. Models including covariates did not converge for boys, due to low base rates of some transitions.


```{r pmat-boys2}
ex_group_boys_2_pmat <- pmatrix.msm(ex_group_boys_2, t = 1, ci = 'normal')

estimate <- as.data.frame(ex_group_boys_2_pmat[['estimates']]) %>% 
  rownames_to_column('From') %>% 
  pivot_longer(c('No EWL', 'EWL', 'Maladaptive Ex'), names_to = 'To', values_to = 'P' ) %>% 
  mutate(P = sprintf("%.3f", as.numeric(P))) 


l <- as.data.frame(ex_group_boys_2_pmat[['L']]) %>%   
  mutate('From' = c('No EWL', 'EWL', 'Maladaptive Ex')) %>% 
  rename('No EWL' = V1) %>% 
  rename('EWL' = V2) %>% 
  rename('Maladaptive Ex' = V3) %>% 
  pivot_longer(c('No EWL', 'EWL', 'Maladaptive Ex'), names_to = 'To', values_to = 'L' ) %>% 
  mutate(L = sprintf("%.3f", as.numeric(L))) 


u <- as.data.frame(ex_group_boys_2_pmat[['U']]) %>% round(3)  %>% 
  mutate('From' = c('No EWL', 'EWL', 'Maladaptive Ex')) %>% 
  rename('No EWL' = V1) %>% 
  rename('EWL' = V2) %>% 
  rename('Maladaptive Ex' = V3) %>% 
  pivot_longer(c('No EWL', 'EWL', 'Maladaptive Ex'), names_to = 'To', values_to = 'U' ) %>% 
  mutate(U = sprintf("%.3f", as.numeric(U))) 

estimate <- cbind(estimate, l$L) 
estimate <- cbind(estimate, u$U)

estimate <- estimate %>% 
  rename('l' = `l$L`) %>% 
  rename('u' = `u$U`)

 
estimate$CI <- paste(estimate$l, estimate$u, sep = ', ')
addparens <- function(x){paste('(',x,')')}
estimate$CI <- addparens(estimate$CI)
estimate$P <- paste(estimate$P, estimate$CI, sep = ' ')

estimate <- estimate %>% 
  select (c('To', 'From', 'P')) %>% 
  pivot_wider(names_from = 'To', values_from = 'P') %>% 
  rename(' ' = From)

knitr::kable(estimate, 
             caption  = 'Estimates of Transition Probabilities Across 1-year amongst boys: From (rows) - To (cols)')
```
Note. EWL = 'Exercise for Weight Loss'. Maladaptive Ex = 'Maladaptive Exercise'


While Markov state models are useful in describing overall transition patterns, one notable limitation is the assumption under these models that transitions are time independent; that is, that that transition to and from states is not dependent on age. This removes consideration of developmental timepoints that may be relevant as risk progresses. Further, the inclusion of covariates is able to test the effect of these covariates on specific transitions, but not overall risk of exercise for weight loss or driven exercise over time. In the next step of analysis, mixed models examine 
