--- 
title: "Associations between Physical Activity Trajectories in Childhood and Driven Exercise in Adolescence in the ALSPAC Cohort"
shorttitle: 'ALSPAC PA Trajectories'
author: 
- name: "Katherine Schaumberg"
  affiliation : '1'
  corresponding: yes
  address: '6001 Research Park Blvd, Madison, WI, USA, 53719'
  role: 
  - Conceptualization
  - Writing - Original Draft Preparation 
  - Writing - Review & Editing
date: "`r Sys.Date()`"
documentclass: 'apa6'
site: bookdown::bookdown_site
bibliography: [PA_trajectories.bib, packages.bib]
biblio-style: apa
link-citations: yes
github-repo: embark-lab/ALSPAC-PA-trajectories
description: "This is a bookdown for ALSPAC PA trajectories Paper"
resource_files: 
output: bookdown::gitbook

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(dplyr)
library(tidyr)
library(tibble)
library(ggplot2)
library(haven)
library(usethis)
library(stringr)
library(rlang)
library(purrr)
library(labelled)
```

<!--chapter:end:index.Rmd-->

# Introduction


max was here
lauren was here <3
Lauren's Intro:

Driven exercise (EDx), or exercise that has a compulsive quality, is a common symptom of eating disorders (ED) (@stiles-shieldsDrivenExerciseTreatmentSeeking2011a). When unable to exercise, individuals who experience EDx feel extreme guilt or anxiety for not exercising and may exercise even when injured or having prior engagements (@stiles-shieldsDrivenExerciseTreatmentSeeking2011a). The typical age of onset for EDx patterns occurs during adolescence, which is also the age when more than 25% of individuals engage in exercise for weight loss (EWL) (@schaumbergProspectiveAssociationsDriven2022a). Due to the overlap in development of EWL and DEx behaviors during late childhood, it is imperative to assess the possible precursor of engagement in physical activity in adolescence to DEx as an eating disorder symptom during early adulthood.

Sarah was here
This is the introduction to the Physical Activities trajectory paper. 

Exercise, which is often associated with positive mental and physical health outcomes in the general population [@penedoExerciseWellbeingReview2005; @chekroudAssociationPhysicalExercise2018; @deslandesExerciseMentalHealth2009], can also become maladaptive, particularly in the context of disordered eating [@dittmerCompulsiveExerciseEating2018; @dallegraveCompulsiveExerciseControl2008]. Though definitions of maladaptive exercise vary and many terms have been used to describe this behavior, a broad conceptualization of maladaptive exercise includes exercise that exhibits one or more of the following qualities: a) exercise that is driven or compulsive in nature, b) feelings of distress or guilt arise when an individual is unable to exercise, c) exercise that interferes with one's ability to engage in school, work, or daily activities, d) the inability to refrain from exercise despite circumstances (e.g. illness, injury) in which it is not recommended [@gorrellPsychosocialEtiologyMaladaptive2021; @scharmerCompulsiveExerciseExercise2020; @dittmerCompulsiveExerciseEating2018]. Such experiences most often arise when individuals also report a primary motivation for engaging in exercise related to weight loss [@scharmerCompulsiveExerciseExercise2020; @zmijewskiExerciseDependenceAttitudes2003]. Most concerning, maladaptive exercise may be an early indicator of disordered eating behavior, as retrospective reports of individuals with eating disorders suggest that maladaptive exercise can be an early-developing eating disorder feature [@stiles-shieldsDrivenExerciseTreatmentseeking2011; @davisRolePhysicalActivity1994]. Further, our previous prospective study in the Avon Longitudinal Study of Parents and Children (ALSPAC) indicated that young people who reported either exercise for weight loss or maladaptive exercise at age 14 were also more likely to report higher concurrent and future (Age 16) eating disorder behaviors [@schaumbergProspectiveAssociationsDriven2022], highlighting the potential etiological relevance of this behavior. Once entrenched, problematic exercise patterns can be a pernicious and debilitating eating disorder symptom and a potential marker of poor prognosis [@brosofExploringContributionExercise2020; @dallegraveCompulsiveExerciseControl2008; @monellRunningEmptyNationwide2018; @solenbergerExerciseEatingDisorders2001]. Amongst individuals with eating disorders, maladaptive (and, more specifically, driven and compulsive) exercise is a commonly-reported feature of illness, with a national clinical eating disorder database estimating that this transdiagnostic symptom is present in approximately half of all eating disorder patients presenting to tretament [@monellRunningEmptyNationwide2018]. 

Given the risks associated with maladaptive exercise patterns and the potential etiologic relevance of exercise behavior the context of eating disorders, clarification regarding both the prevalence of and early predictors of high-risk exercise behavior during adolescence could build knowledge necessary to prevent maladaptive exercise patterns from taking hold. Previous research highlights several potential predictors of maladaptive exercise in young people, including, at a social level: pressures to conform to body image ideals [@reynoldsSocioculturalInfluencesCompulsive2022], at a psychological level: weight and shape concerns, negative affect, thinnness expectations [@gorrellPsychosocialEtiologyMaladaptive2021], and, at a biolgoical level: genetic risk for obsessive-compulsive disorder and/or anorexia nervosa [@yilmazPredictingEatingDisorder2022].

Altogether, initial evidence suggests that engaging in exercise for weight loss could escalate to maladaptive exercise and trigger eating disorder risk in some cases, though additional investigation is necessary to clarify the processes that influence exercise-related risk and the frequency with which young people transition to and from maladaptive patterns of exercise. Further, identification of early predictors of maladaptive exercise could aid in identifying youth at risk for the development of disordered eating.

## Current Study

In the current study, we extend a longitudinal investigation of exercise for weight loss and maladaptive exercise across a developmental window from middle adolescence to young adulthood (ages 14-24). Specifically, we aim to examine the frequency of exercise states, based on the three categories of 'no exercise for weight loss', 'exercise for weight loss', and 'maladaptive exercise', amongst boys and girls at ages 14, 16, 18, and 24 years old. We examine rates of transitions between these exercise states over time, and identify preliminary predictors of these transitions. Further, we aim to characterize overall risk for engaging in maladaptive exercise across age, and identify potential demographic or early (age 14) eating disorder cognition variables that may enhance risk for maladaptive exercise throughout adolescence. Based on previous literature [@allenEatingDisorderSymptom2013; @schaumbergProspectiveAssociationsDriven2022; @davisDevelopmentalTrajectoriesCompensatory2016], we hypothesize that exercise for weight loss will be a common behaivor across adolescence, with maladaptive exercise being less common than exercise for weight loss, but will occur in a sizable minority of individuals. Further, we hypothesize that eating disorder cognitions assessed at age 14 will be associated with higher likelihood of exercise for weight loss and maladaptive exercise across adolescence. Specific aims are as follows:


Lauren's Intro:

Driven exercise (EDx), or exercise that has a compulsive quality, is a common symptom of eating disorders (ED) [@stiles-shieldsDrivenExerciseTreatmentSeeking2011]. When unable to exercise, individuals who experience EDx feel extreme guilt or anxiety for not exercising and may exercise even when injured or having prior engagements [@stiles-shieldsDrivenExerciseTreatmentSeeking2011]. The typical age of onset for EDx patterns occurs during adolescence, which is also the age when more than 25% of individuals engage in exercise for weight loss (EWL) [@schaumbergProspectiveAssociationsDriven2022]. Due to the overlap in development of EWL and DEx behaviors during late childhood, it is imperative to assess the possible precursor of engagement in physical activity in adolescence to DEx as an eating disorder symptom during early adulthood.


Max Intro:
While previous studies have shown how inactivity in adolescence can lead to negative outcomes later in life[@kannYouthRiskBehavior1996] . Less is known about the affects of maladaptive exeercise in adolesence and how it relates to adulthood. Analysis of the Avon Logitudinal Study of Parents and Children Cohort (ALSPAC) in adolesence and adulthood will help to establish the connection beweeen adolescent and adulthood physical activity [@boydCohortProfileChildren2013] [@fraserCohortProfileAvon2013] 

Sarah Intro:
The Avon Longitudinal Study of Parents and Children (ALSPAC), a transgenerational observational study investigating influences on health across the life course includes data describing childhood and adult activity levels [@boydCohortProfileChildren2013]. Activity levels in early life are thought to influence adult levels of activity.... 


<!--chapter:end:01-Introduction.Rmd-->


# Method

## Sample - The ALSPAC Cohort


The ALSPAC Cohort [@boydCohortProfileChildren2013; @fraserCohortProfileAvon2013] was established to understand how genetic and environmental characteristics influence health and development in parents and children. Ethical approval for this study was granted by the ALSPAC Law and Ethics Committee and Local Ethics Committees. All pregnant women living in the geographical area of Avon, United Kingdom, who were expected to deliver between April 1, 1991 and December 31, 1992, were invited to participate in the study. Children from 14,541 pregnancies were enrolled; 13,988 children were alive at 1 year. An additional 913 children were enrolled during subsequent phases of enrollment, with a total sample size alive at 1 year of 14,901. All women gave informed and written consent. Among twin pairs, one twin per pair was randomly excluded from the current study.

A fully searchable ALSPAC data dictionary is available [here](http://www.bris.ac.uk/alspac/researchers/data-access/data-dictionary/).


11,13,15 - PA

14,16,18,24 - DEx

Inclusion - have to have one of each PA and DEx


<!--chapter:end:02-Method_Sample.Rmd-->

## Measures

### Socioeconomic Status (SES)

Measures of parent occupation were obtained for both mother and partner during child gestation. Parental occupations were then classified as unskilled(0), partly skilled (1), skilled manual (2), skilled non-manual (3), manegerial and technical (4), or professional (5). Of reported parent occupations, the highest reported occupation was used as a proxy for parent SES, to be treated as a continuous (0-5) covariate in analyses.

### Body Mass Index Z-score (BMI-Z) at Age 10

```{r bmi, include = FALSE}
# clinic <- sum(!is.na(demo_ed.f$bmi_clinic.13))
# no_clinic <- sum(is.na(demo_ed.f$bmi_clinic.13))
# best_avail <- sum(!is.na(demo_ed.f$bmi_bestavail.13))
# 
# missing_n <- 7688- best_avail
# pr_n <- best_avail-clinic
# missing_percent <- round ((7688 - best_avail)/7688*100 , 2)
# clinic_percent <-  round ((clinic)/7688*100, 2)
# pr_percent <- round((best_avail-clinic)/7688*100, 2)

```

Child BMI-Z was assessed via two indices at age 13 years. First, objective height and weight were obtained for those who completed a clinic visit at age 13 (median age at clinic visit = 13.8 years). Second, parents reported child height and weight via questionnaire (median age = 13.1). An age and sex adjusted BMI-Z score was obtained from these self-report and clinic-recorded measurments. In order to maximize accuracy, when clinic-obtained measurements were obtained, this BMI-Z score was used (n = r clinic; clinic_percent %. BMI-Z from parent report was used when clinic measurement was not available (n = pr_n; pr_percent%). A portion of the sample had neither measurement available (n = missing_n\`; missing_percent%)

### Exercise Measures

Questions related to maladaptive exercise were self-reported at ages 14, 16, 18, and 24 years. Questions were adapted from the Youth Risk Behavior Survelliance System [@kannYouthRiskBehavior1995].

#### Exercise for Weight Loss

At ages 14, 16, and 18 years, participants were asked if they exercised to lose weight or avoid gaining weight, with response options of: No(0), Yes-Sometimes(1), and Yes-Frequently(2).

At age 24, participants were asked the frequency with which they exercised to lose weight or avoid gaining weight, with response options of: Never, \<1x/mo, 1-3x/mo, 1-4x/wk, and 5 or more times per week. For the purposes of analysis, responses at age 24 were harmonized with age 14, 16, and 18. Responses of Never were binned as "No" (0); 1x/mo - \<1x/week were binned as 'Yes - Sometimes'(1), and 1x/wk or more as 'Yes - Frequently'(2). Once a week or more was chosen as the threshold for 'frequent' exercise for weight loss at age 24 to align with DSM-V diagnostic frequency thresholds for other eating disorder behaviors, such as binge eating and purging [@americanpsychiatricassociationDiagnosticStatisticalManual2013]. The primary derived outcome variable of analysis for exercise for weight loss is an ordinal variable, defined as 0 = No, 1 = Sometimes, and 2 = Frequently.

#### Maladaptive Exercise

Issues that are associated with exercise were also reported at ages 14,16, 18, and 24 years. At all assessment points, participants reported whether exercise interferes with work/school (14,16,18), or their daily routine (24). At ages 14 and 24, participants also reported whether they exercised to lose weight even when sick or injured. At ages 16 and 18 years, participants reported whether they felt guilty about missing an exercise session -- see Table \@ref(tab:tab2) . Responses options for exercise issues were 'No', 'yes - Sometimes', or 'Yes - Frequently'.

```{r tab2}
# x <- tibble(age = c(14,16,18,24), Interferes = c('x','x','x','x'), Sick_Injured = c('x',' ',' ', 'x'), Guilt = c('' ,'x','x', '')) 
# 
# knitr::kable(x, caption = 'Exercise Issues Assessed at Each Age')
```

Exercise issues are deemmed to be 'present' at each time point if an individual endorses: (1) exercise interfering with work/school sometimes or more OR frequent guilt when missing an exercise session at ages 16 and 18 years and (2) exercise interfering with work/school/daily routine sometimes or more OR exercising even when sick/injured sometimes or more at ages 14 and 24 years.

The presence of maladaptive exercise in this study was defined as exercising for weight loss **sometimes or frequently** and at least one **exercise issue** defined as present at each time point. Frequency of exercise for weight loss as sometimes or more was chosen for this definition as (1) existing literature suggests that frequency of exercise may be less important than cognitive features in defining exercise risk in the context of eating disorders [@adkinsDoesExcessiveCompulsive2005; @mondUpdateDefinitionExcessive2006] and (2) the nature of the epidemiological sample and the concomitant goals of this paper are such that we are interested in identifying early risk indicators, which may be best captured via a broad (vs. narrow) definition of maladaptive exercise. The primary variable to define maladaptive exercise was derived as described, resulting in a dichotomous outcome (0 = Absent, 1 = Present).

#### Exercise Gropus

Exericse groups over time are defined for descriptive analyses and transition states, replicating groups defined in @schaumbergProspectiveAssociationsDriven2022. Those who report no exericse for weight loss at each time point are placed in a 'No Exercise for Weight Loss' group, those who report exercise for weight loss at least sometimes, but do not meet criteria for maladaptive exercise are placed in an 'Exercise for Weight Loss' group, and those meeting maladaptive exercise criteria are placed in the 'Maladaptive Exercise' group.

#### Eating Disorder Behaviors

The measurement of eating disorder behaviors in adolescence captures a more complete range of behaviors beyond ED diagnosis. Eating disorder behaviors were assessed through questionnaires developed from the Youth Risk Behavior Surveillance System (YRBSS) at Waves 14+, 16+, 18+, and 24+ years. Participants were asked to consider the past year in their responses.

| Construct \| Timepoint \| Description \| \| \|\|-------\|-------\|----------------------------------------------\|-------\|-------\|
| Thin ideal internalization \| 14 \| Thin ideal internalization was measured at age 14 with the Ideal_Body Stereotype Scale_Revised (Stice, 1998). Adolescents responded to statements regarding appearance ideals rated on a five_point Likert scale from 'strongly agree' to 'strongly disagree.' \| \| \|
| Body Satisfaction \| 14, 18, 24 \| The Body Satisfaction Scale was measured at 14, 18, and 24 and asked individuals to rate their satisfaction with their body parts on a 5-point Likert scale, from 'extremely satisfied' to 'extremely dissatisfied' with higher values indicating higher dissatisfaction. \| \| \|
| Pressure to lose weight \| 14 \| Pressure to lose weight was measured at age 14 with an adapted version of the Perceived Sociocultural Pressure Scale. Pressure to lose weight from friends, family, romantic partners, and the media was recorded. \| \| \|
| Pressure to increase muscle \| 14 \| Pressure to gain muscle was measured at age 14 with an adapted version of the Perceived Sociocultural Pressure Scale. Pressure to gain muscle from friends, family, romantic partners, and the media was recorded. \| \| \|
| Weight-related teasing \| . \| Adolescents were asked to report weight related teasing from peers and family members with five possible responses ranging from "never" to "at least once a week." \| \| \|
| Dieting \| 14, 16, 18 \| Participant dieting was measured at age 14, 16, and 18 and measured frequency during the past year YP went on a diet to lose weight/keep from gaining weight, length of diet, amount of weight lost, and amount of weight regained. \| \| \|
| External Eating \| 14 \| Participants completed a variation of the Dutch Eating Behavior Questionnaire (DEBQ) at age 14. Dysfunctional eating patterns such as emotional eating, external eating, and restrained eating were assessed. \| \| \|
| Restrained Eating \| 14 \| Participants completed a variation of the Dutch Eating Behavior Questionnaire (DEBQ) at age 14. Dysfunctional eating patterns such as emotional eating, external eating, and restrained eating were assessed. \| \| \|
| Emotional Eating \| 14 \| Participants completed a variation of the Dutch Eating Behavior Questionnaire (DEBQ) at age 14. Dysfunctional eating patterns such as emotional eating, external eating, and restrained eating were assessed. \| \| \|
| Fear of Weight gain \| 14, 24 \| Participants asked the degree they worried about gaining a little weight (two pounds) in the past year. \| \| \|
| Maternal Dieting \| 14 \| Mothers were asked the frequency they tried to lose weight in the past year. \| \| \|
| Craving \| 16 \| Participants were asked the frequency of occurrence of strong cravings for food that they find hard to resist. \| \| \|
| Weight intentions \| 14, 18 \| Respondent asked about what they are trying to do with their weight. \| \| \|
| Weight Perception \| 14, 18 \| Participant notes their perception of their weight. \| \| \|
| Weight valuation \| 18 \| Participant notes the degree to which their weight has made a difference to how they feel about themselves. \| \| \|
| Feeling fat \| 24 \| Participant notes the frequency that they have felt fat in the past year. \| \| \|
| Weight/Shape concern \| 14 \| Weight and shape concern at age 14 was assessed using a 4 item questionnaire adapted from the McKnight Risk Factor Survey. \| \| \|

### Physical Activity

#### Accelerometers

Physical activity data was captured using the Actigraph^tm^ accelerometer, which is often used in child physical activity studies. Participants were instructed to wear accelerometers around the waist just above their right hip for the duration of the tracking period. Accelerometer data was collected at the following four time points. At ages 11.5, 13.5, and 15.5, during a clinic visit, young people were asked to wear the accelerometer for 7 days, beginning the day after clinic attendance [@mattocksUseAccelerometersLarge2008]. At age 24 participants were asked to wear accelerometers for four days beginning the day after clinic attendance [@troianoPhysicalActivityUnited2008].

#### CPM

Accelerometer data was used to extract total physical activity which was measured as the total counts per minute (CPM) over the period of valid recording. Associations with total physical activity are usually calculated per 100 cpm as this difference is of a similar order to the differences observed between males and females in these data.

#### MVPA

Moderate to Vigerous Physical Activity (MVPA) measures the average amount of time spent above a CPM cutoff, these cutoff points differ by age groups. At ages 11.5, 13.5 and 15.5 A value of \>= 3600 cpm was originally used to define MVPA at the time of the data collection [@mattocksUseAccelerometersLarge2008]. More recently, following Evenson et. al. [@evensonCalibrationTwoObjective2008], who calculated a lower cutpoint of \>=2296. Both sets of data were included in analysis. At age 24 MVPA cutoff was established at \>2020 cpm [@troianoPhysicalActivityUnited2008].

#### Valid Days

Valid Days are calculated using a cutoff of minutes per day that an accelerometer was worn. For ages 11.5-15.5 a cutoff of \>=600 minutes per day was used. At age 24 Valid Days were operationalised as a wear time of at least 500 minutes, after excluding intervals of \>60 minutes with zero counts. \#### Sedentary Behavior

### Eating Disorder Diagnoses

Eating disorder diagnoses (AN, BN,BED, purging disorder [PD], and Other Specified Feeding and Eating Disorder (OSFED)) for adolescents were derived using questionnaire data from the Youth Risk Behavior Surveillance System questionnaire (YRBSS) using DSM-5 diagnostic criteria (excluding OSFED) [@micaliAdolescentEatingDisorders2015][@americanpsychiatricassociationDiagnosticStatisticalManual2013]. AN was assessed using both the YRBSS self-report and parental report since adolescents tend to under-report AN symptoms [@houseDiagnosingEatingDisorders2008]. Table X shows the criteria used for each diagnosis [@micaliAdolescentEatingDisorders2015].

Diagnoses were assessed at Waves14+, 16+, 18+, and 24+. Body mass index (BMI), which is an objective measure, was also collected at face to face assessment (median ages: 13.8, 15.5, 17.8, and 24.0) and used as a diagnostic criteria for AN. Furthermore, an underweight assignment wasdetermined using age, gender, and BMI-specific cutoffs (based on UK reference data); corresponding to World Health Organization (WHO) grade 1 thinness [@coleBodyMassIndex2007]. Excessive exercise, which is used as criteria for an AN and OSFED diagnoses, was assigned to the adolescent if they found themselves lacking time to do schoolwork due to engaging in exercise, exercising when sick or injured, and feeling guilt when unable to exercise.

| **Table X. Diagnostic Criteria**                                    |                 |                                                                                                                                                                                                              |                                                                                                                                                                                                              |                                                                                                                                                                                                              |                                                                                    |
|------------|------------|------------|------------|------------|------------|
| **Diagnoses**                                                       | **Report Type** | **Wave 14+**                                                                                                                                                                                                 | **Wave 16+**                                                                                                                                                                                                 | **Wave 18+**                                                                                                                                                                                                 | **Wave 24+**                                                                       |
| Anorexia Nervosa (AN)                                               | Self-report     | 1.Underweight (body mass index [BMI]\<18.5)                                                                                                                                                                  | 1.Underweight (BMI\<18.5)                                                                                                                                                                                    | 1\. engaged in fasting for weight loss or to avoid weight gain at least monthly                                                                                                                              | 1.Underweight (BMI\<18.5)                                                          |
|                                                                     |                 | AND 2. self-reported weight and shape concern                                                                                                                                                                | AND 2. engaged in fasting for weight loss or to avoid weight gain at least monthly                                                                                                                           | OR engaged in excessive exercise                                                                                                                                                                             | AND 2. engaged in fasting for weight loss or to avoid weight gain at least monthly |
|                                                                     |                 | OR engaged in fasting for weight loss or to avoid weight gain at least monthly                                                                                                                               | OR engaged in excessive exercise                                                                                                                                                                             |                                                                                                                                                                                                              | OR engaged in excessive exercise                                                   |
|                                                                     |                 | OR engaged in excessive exercise                                                                                                                                                                             |                                                                                                                                                                                                              |                                                                                                                                                                                                              |                                                                                    |
|                                                                     | Parental Report | 1\. Underweight (BMI\<18.5)                                                                                                                                                                                  | 1\. Underweight (BMI\<18.5)                                                                                                                                                                                  |                                                                                                                                                                                                              |                                                                                    |
|                                                                     |                 | AND 2. Presence of fear of weight gain AND fat avoidance in the 3 months prior to assessment                                                                                                                 | 2\. Presence of fear of weight gain AND fat avoidance in the 3 months prior to assessment                                                                                                                    |                                                                                                                                                                                                              |                                                                                    |
| Bulimia Nervosa (BN)                                                | Self-Report     | 1.Weekly binge eating                                                                                                                                                                                        | 1.Weekly binge eating                                                                                                                                                                                        | 1.Weekly binge eating                                                                                                                                                                                        | 1.Weekly binge eating                                                              |
|                                                                     |                 | AND 2. weekly purging                                                                                                                                                                                        | AND 2. weekly purging                                                                                                                                                                                        | AND 2. weekly purging                                                                                                                                                                                        | AND 2. weekly purging                                                              |
| Binge Eating Disorder (BED)                                         | Self-Report     | 1\. Weekly binge eating                                                                                                                                                                                      | 1.Weekly binge eating                                                                                                                                                                                        | 1.Weekly binge eating                                                                                                                                                                                        | 1\. Weekly binge eating                                                            |
|                                                                     |                 | AND 2. At least 3 cognitive symptoms (eating fast or faster than normal; eating until stomach hurt or they felt sick, eating large amounts when not hungry, eating alone, feeling guilty about amount eaten) | AND 2. At least 3 cognitive symptoms (eating fast or faster than normal; eating until stomach hurt or they felt sick, eating large amounts when not hungry, eating alone, feeling guilty about amount eaten) | AND 2. At least 3 cognitive symptoms (eating fast or faster than normal; eating until stomach hurt or they felt sick, eating large amounts when not hungry, eating alone, feeling guilty about amount eaten) | AND 2. Absence of purging                                                          |
|                                                                     |                 | AND 3. Absence of purging                                                                                                                                                                                    | AND 3. Absence of purging                                                                                                                                                                                    | AND 3. Absence of purging                                                                                                                                                                                    |                                                                                    |
| Purging Disorder                                                    | Self-Report     | 1\. Purging at least weekly                                                                                                                                                                                  | 1\. Purging at least weekly                                                                                                                                                                                  | 1\. Purging at least weekly                                                                                                                                                                                  | 1\. Purging at least weekly                                                        |
|                                                                     |                 | 1\. AND 2. Binge eating absent or \< monthly                                                                                                                                                                 | AND 2. Binge eating absent or \< monthly                                                                                                                                                                     | AND 2. Binge eating absent or \< monthly                                                                                                                                                                     | AND 2. Binge eating absent or \< monthly                                           |
| Sub-threshold BN (Sub-BN)                                           | Self-Report     | 1\. monthly binge eating                                                                                                                                                                                     | 1\. monthly binge eating                                                                                                                                                                                     | 1\. monthly binge eating                                                                                                                                                                                     | 1\. monthly binge eating                                                           |
|                                                                     |                 | AND 2. monthly purging                                                                                                                                                                                       | AND 2. monthly purging                                                                                                                                                                                       | AND 2. monthly purging                                                                                                                                                                                       | AND 2. monthly purging                                                             |
| Sub-threshold BED (Sub-BED)                                         | Self-Report     | 1\. monthly binge eating                                                                                                                                                                                     | 1\. monthly binge eating                                                                                                                                                                                     | 1\. monthly binge eating                                                                                                                                                                                     | 1\. monthly binge eating                                                           |
|                                                                     |                 | AND 2. At least 3 cognitive symptoms                                                                                                                                                                         | AND 2. At least 3 cognitive symptoms                                                                                                                                                                         | AND 2. At least 3 cognitive symptoms                                                                                                                                                                         | AND 2. Absence of purging                                                          |
|                                                                     |                 | AND 3. Absence of purging                                                                                                                                                                                    | AND 3. Absence of purging                                                                                                                                                                                    | AND 3. Absence of purging                                                                                                                                                                                    |                                                                                    |
| Other Specified Feeding and Eating Disorder-other 1 (OSFED-other 1) | Self-Report     | monthly binge eating, purging, excessive exercising or fasting                                                                                                                                               | monthly binge eating, purging, excessive exercising or fasting                                                                                                                                               | monthly binge eating, purging, excessive exercising or fasting                                                                                                                                               | monthly binge eating, purging, excessive exercising or fasting                     |
| Other Specified Feeding and Eating Disorder-other 2 (OSFED-other 2) | Self-Report     | \< monthly binge eating, purging, excessive exercise or fasting                                                                                                                                              | \< monthly binge eating, purging, excessive exercise or fasting                                                                                                                                              | \< monthly binge eating, purging, excessive exercise or fasting                                                                                                                                              | \< monthly binge eating, purging, excessive exercise or fasting                    |
|                                                                     |                 | AND weight and shape concern                                                                                                                                                                                 |                                                                                                                                                                                                              |                                                                                                                                                                                                              |                                                                                    |

<!--chapter:end:03-Method_Measures.Rmd-->


# Analytic Plan

Analyses will be gender stratified.
 

<!--chapter:end:04-Analytic_Plan.Rmd-->


# Results

## Descriptives

### Covariates: SES, BMIZ at Age 13, ED Cognitions at age 14

Descriptive information for continuous variables are provided in Supplemental Table \@ref(tab:continuous). Descriptives for ordinal variables (Parent SES and Fear of Weight Gain) are provided in Supplemental Tables \@ref(tab:pses) and \@ref(tab:fearwt)

### Exercise for Weight Loss Frequency

Exercise for weight loss frequency across age and gender is presented in Figure \@ref(fig:ewl-plot), with a plot of the raw frequencies at age 24 presented in Supplement (Figure \@ref(fig:age24-raw-ewl)). Visual inspection of plots suggests a potential increase in frequency of exercise for weight loss across age, with a consistent increase over time amongst girls and an increase in frequency particularly between ages 18 and 24 years amongst boys.


### Exercise Issues Frequency

With regards to exercise issues across time, the percentage of participants reporting any exercise issues as present at each age was relatively stable, as seen in Figure \@ref(fig:ex-issues). Although there was some variation in reporting of the secondary question regarding exercise issues across age which could influence the detection of exercise issues at different ages, the question of whether exercise interfered with work, school, or one's daily routine was assessed at all four timepoints. Endorsement of this question, specifically, was relatively consistent across age and gender (See Supplemental Figure \@ref(fig:ex-interfere)), with 9.5 - 14.5% of individuals reporting that exercise interfered with work, school, or their daily routine across age and gender. 

### Exercise Groups

Figure \@ref(fig:freqfig) shows exercise status at ages 14-24 years in girls and boys, as a percentage of the total number of boys and girls who reported exercise data at the timepoint.

```{r freqfig, fig.cap = 'Frequency of Exercise Groups Across Age'}
# 
# exercise_group_data <- frq_table_by_age_sex(demo_ed.lf, exercise_group) 
# levels(exercise_group_data$exercise_group) <- c('No Exercise for Weight Loss', 'Exercise for Weight Loss', 'Maladaptive Exercise')
# 
# exercise_status_plot_1 <- ggplot(data = exercise_group_data, 
#     aes(x = sex, y = pct, fill = exercise_group, label = sprintf('%0.1f%%', pct*100))) + #set the aesthetic features, including the labels at one decimal place
#     geom_bar(position = position_stack(reverse = TRUE), stat = 'identity') + #add column graph, with male and female separated
#     scale_y_continuous(labels = scales::percent, limits = c(0,1))+ #set the lower and upper limits of the y axis ( 0-100 percent); 
#     theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), text = element_text(size = 15))+ #make legend text a bit bigger
#     labs(x = element_blank(), y = 'Percent (within gender)', title = 'Exercise Groups Across Age') +
#      geom_text(aes(label = sprintf('%0.1f%%', pct*100)), position = position_stack(reverse = TRUE, vjust = 0.5), color = 'white', size = 2.5, fontface = 'bold') +
#     facet_wrap(~age) + 
#     theme_classic()+
#     theme(legend.title = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1), text = element_text(size = 15)) + #make legend text a bit bigger 
#    scale_fill_manual(values = wes_palette(name = 'GrandBudapest1', n = 3 ))
# 
# exercise_status_plot_1
```

Visual inspection and raw percentage comparison indicates that percentage of males and female in the three exercise groups is relatively stable from ages 14-16 years, with a possible increase in exercise for weight loss at age 18 among girls, an increase in exercise for weight loss among men at age 24, and an increase in maladaptive exercise across gender at age 24.



<!--chapter:end:05-Descriptives.Rmd-->

# Discussion

<!--chapter:end:06-discussion.Rmd-->


# References {-}

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```


`r if (knitr:::is_html_output()) '

'`

<!--chapter:end:07-References.Rmd-->

