
#Mixed Effects Models

## Driven Exercise

### Girls

```{r}
load('Girl_Models')
```

The below analyses (1) compare the baseline random effects model with
the random intercept model and (2) compare the random intercept model describe the fixed effects for age



```{r}

bl_0_dex_girls_pooled <- pool(bl_0_dex_girls)
bl_0_dex_girls_ORs <- summary(bl_0_dex_girls_pooled, conf.int = TRUE, exponentiate = TRUE)

age_1_dex_girls_pooled <- pool(age_1_dex_girls)
age_1_dex_girls_ORs <- summary(age_1_dex_girls_pooled, conf.int = TRUE, exponentiate = TRUE)

#use the D1 method to compare 
mcomp_1 <- D1(age_1_dex_girls, bl_0_dex_girls)

```

This model indicates that there are increasing odds of driven exercise
over time, with $e^{-2.75}$ at age 14 indicating low average predicted
odds of driven exercise at age 14 (6.4%), and average increase across
age with an odds ratio of $e^{0.12}$ = 1.13 each year. The random
intercept model outperformed the random effects model.

```{r}

#Now we include fixed effect covariates of BMI, parent highest occupation, and eating disorder cogitions at age 14: thin-ideal internalization, body dissatisfaction, and fear of weight gain
covs_2_dex_girls_pooled <- pool(covs_2_dex_girls)
covs_2_dex_girls_ORs <- summary(covs_2_dex_girls_pooled, conf.int = TRUE, exponentiate = TRUE)

mcomp_2 <- D1(covs_2_dex_girls, age_1_dex_girls)


covs_3_dex_girls_pooled <- pool(covs_3_dex_girls)
covs_3_dex_girls_ORs <- summary(covs_3_dex_girls_pooled, conf.int = TRUE, exponentiate = TRUE)

mcomp_3 <- D1(covs_3_dex_girls, covs_2_dex_girls)

covs_4_dex_girls_pooled <- pool(covs_4_dex_girls)
covs_4_dex_girls_ORs <- summary(covs_4_dex_girls_pooled, conf.int = TRUE, exponentiate = TRUE)

mcomp_4 <- D1(covs_4_dex_girls, covs_3_dex_girls)

```


```{r, label = 'DEx Poled Coeffs Table - Girls'}

dfnames <- c("bl_0_dex_girls_ORs", "age_1_dex_girls_ORs", 'covs_2_dex_girls_ORs', 'covs_3_dex_girls_ORs', 'covs_4_dex_girls_ORs')
pooled_coefs <- do.call(rbind, lapply(dfnames, function(x) cbind(get(x), Model=x))) |> 
  relocate(Model) %>% 
  select(-c(p.value, statistic, df))

for (i in 3:ncol(pooled_coefs) ) {
  pooled_coefs[i] <- round(pooled_coefs[i], digits = 3)
}

pooled_coefs['Model'][pooled_coefs['Model'] == 'bl_0_dex_girls_ORs'] <- 'Baseline Model'
pooled_coefs['Model'][pooled_coefs['Model'] == 'age_1_dex_girls_ORs'] <- 'Age Effect Model'
pooled_coefs['Model'][pooled_coefs['Model'] == 'covs_2_dex_girls_ORs'] <- 'Step 1 Covariates'
pooled_coefs['Model'][pooled_coefs['Model'] == 'covs_3_dex_girls_ORs'] <- 'Step 2 Covariates'
pooled_coefs['Model'][pooled_coefs['Model'] == 'covs_4_dex_girls_ORs'] <- 'Step 3 Age x Cov Interactions'

pooled_coefs[] <- lapply(pooled_coefs, as.character)
pooled_coefs['term'][pooled_coefs['term'] == 'parent_highest_occupation'] <- 'Parent SES'
pooled_coefs['term'][pooled_coefs['term'] == 'bmi_z_bestavail'] <- 'BMI Z - Age 13'
pooled_coefs['term'][pooled_coefs['term'] == 'age_adjust'] <- 'Age'
pooled_coefs['term'][pooled_coefs['term'] == 'fear_wtgain_14'] <- 'Fear of Wt Gain - Age 14'
pooled_coefs['term'][pooled_coefs['term'] == 'body_sat_mean_14_std'] <- 'Body Satisifaction - Age 14'
pooled_coefs['term'][pooled_coefs['term'] == 'ti_mean_14_std'] <- 'Thin-ideal Internalization - Age 14'
pooled_coefs['term'][pooled_coefs['term'] == 'age_adjust:bmi_z_bestavail'] <- 'Age x BMI (13)'
pooled_coefs['term'][pooled_coefs['term'] == 'age_adjust:fear_wtgain_14'] <- 'Age x Fear of Wt Gain (14)'
pooled_coefs['term'][pooled_coefs['term'] == 'age_adjust:body_sat_mean_14_std'] <- 'Age x Body Satisfaction (14)'
pooled_coefs['term'][pooled_coefs['term'] == 'age_adjust:ti_mean_14_std'] <- 'Age x Thin-Ideal Internalization (14)'


pooled_coefs|> 
  kbl(caption = 'Table 1: Parameter Estimates for Models Predicting Driven Exercise in Girls') %>% 
  row_spec(c(2:3, 8:14), background = '#E0E0E0') %>% 
  kable_paper(full_width = T) 

```
```{r}
#Results Table - Model Comparisons

model_comparisons <- c('mcomp_1', 'mcomp_2', 'mcomp_3', 'mcomp_4')

n = list() 
for (i in model_comparisons) { 
  x<- as.list((eval(ensym(i))[2])) 
  x<- data.frame(t(unlist(x)))
  colnames(x) = c('statistic', 'df1', 'df2', 'p.value','riv')
  n[[i]] = x
}
pooled_coefs_2 <- do.call('rbind', n) 

rownames(pooled_coefs_2) <- c('Baseline vs. Age Effects', 'Age vs. Step 1 Covariates', 'Step 1 vs. Step 2 Covariates', 'Step 2 vs. Age x Covariate Interactions')

pooled_coefs_2 %>% 
  kbl(caption = 'Table 2: Model Comparisons - Driven Exercise in Girls', digits = 3) %>% 
  kable_material('striped')


```

### Boys

Below we compute the same Random Effects and Random Intercept Model with
a Fixed Effect of Age in Boys

```{r}

```

For boys, we see very low average predicted odds of driven exercise
expected at age 14 (0.01%), with odds ratio of 1.32 indicating increased
risk for each year of age after 14. The random intercept model
outpreformed the random effects model.

```{r}
#Now we include fixed effects of: BMI at age 14, Parent SES, thin-ideal internalization at age 14, and body dissatisfaction at age 14

# rand_int_dex_f <- glmer(driven_exercise ~ 1 + age_adjust + BMI_14 + education_mum_d + (1|id), data = demo_ed.l.girls, family = binomial)

#Finally, we examine whether there is an interaction between thin-ideal interanlization at age 14 and age and body dissatisfaction at age 14 and age, such that individuals who exhibit eating disorder cognitions at age 14 are more likely to shift towards driven exercise engagement with age. 

```

## Exercise for Weight Loss

We are utilizing the same approach to examine endorsement of exercise
for weight loss over time in the sample, but use an ordinal model (clmm)
.

<https://rdrr.io/github/DaanNieboer/ordinalimputation/man/pooling.clmm.html>

### Girls

### Boys

m1 <- as.list(mcomp_rand_int_x_bl$result) 
m1 <- data.frame(t(unlist(m1)))

m2 <- as.list(mcomp_cov1_x_rand_int$result)
dfnames<- c('m1', 'm2')
pooled_coefs_1 <- do.call(rbind, lapply(dfnames, function(x) cbind(get(x), Model=x))) 

